Nuclei Segmentation in Histopathology images using Deep Neural Networks

### 2017-01-20 Paper.tex
shortened to 4 pages

\noindent Many traditional techniques have been proposed for the segmentation of
nuclei in histopathology images, ranging from simple background
subtraction and color threshold techniques 
to much more
sophisticated approaches, such as marked point
processes\cite{Kulikova2012}. Many of these methods have been recently
reviewed \cite{irshad2014methods}.

@@

Alternative methods 
have arisen from the achieving Convolutional neural networks.
%\cite{lecun}, 
%in particular strong incentive rose with the very impressive 
%results that they achieved in the pascal voc challenge with 
%\cite{ImageNet}. 
Recent advances in deep neural networks, and in 
particular in their optimization have made them become the state-of-the-
art model for object recognition.  Deep neural network have also been 
used for 
semantic segmentation, where \cite{long2015fcn} use "de-
convolution layers" and up-sampling in order to identify and precisely 
locate objects within a picture. Different architecture arising from 
different intuition are also possible and have been applied in this paper.

Learning based methods rely on annotated data sets. But for the
problem of nuclei segmentation in histopathology images, there are
only few manually segmented images freely available. 
In \cite{Drelie08-298}, a partially annotated data set has been
released. However, this data set does not cover the morphological
variability typically encountered in real histopathology data. Moreover, nuclei spot detection in histopathology has been tackled in breast and colon cancer in \cite{xu2016stacked} and \cite{sirinukunwattana2016locality}. Our setup is different in terms of architecture as we use a very deep architecture for segmentation and not image patches for nuclei centroid detection.

\begin{figure}[htb]
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_1}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 498959}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_1}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
%\begin{minipage}[b]{.48\linewidth}
 % \centering
%  \centerline{\includegraphics[width=4.0cm]{RGB_2}}
%  \vspace{1.5cm}
%  \centerline{Sample from patient 581910}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{0.48\linewidth}
%  \centering
%  \centerline{\includegraphics[width=4.0cm]{GT_2}}
%  \vspace{1.5cm}
%  \centerline{Associated ground truth}\medskip
%\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_3}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_3}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
\caption{Random annotated samples from the dataset}
\label{fig:annotation}
%
\end{figure}

@@

%% Validation scheme
We have our training set $(A_l,B_l)_{l \in [1, N]}$ where N is equal to 
33, however each element $A_l$ belongs to a certain patients and 
several elements $A_l$ can belong to the same patient. As we are 
dealing with histopathology images, it is known that samples can widely 
vary from one patient to the other. We thus validate our model by a 
leave one patient out scheme. 
For this reason we use a leave-one-patient-out strategy: for a 
given set of hyper parameters, we 
train our model on every patient except one that is used for validation. 
Our final score is averaged over all patients.
Several metrics give a detailed assessment of the model quality: accuracy, intersection over union (IU),
F1 score and a performance score which is the mean between 
true positive rate and true negative rate.

@@

%
\caption{Prediction via different classifiers of a random sample on the left out patient: 581910}
\label{fig:prediction}
%
\end{figure}

@@

\vspace{-1cm}
\section{Different architectures}
\label{sec:results}
\noindent We experience with 3 known arhitectures in semantic segmentation, named PangNet, Fully Convolutionnal Net (FCN) and DeconvNet. The most basic architecture, PangNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map, for more information please refer to \cite{pang2010cell}. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
architecture has the advantage of re-using a classical deep learning 
architecture with additional upsampling and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture\cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layers as one intends to learn 
the proper upsampling through repeated deconvolution and convolution 
layers. This model can also be fine tuned\cite{noh2015learning}. We also did an
ensemble classifier of these two previous nets as suggested in 
\cite{noh2015learning} but did not include any illustration. Processing 
one image of size $224 \times 224$ pixels took on average 0.91 secondes for 
PangNet, which is the smallest network, and it took on average 3.38 secondes for DeconvNet, the 
largest network. These times do not depend on the number of cells. These computation 
times were estimated on a K620 Quadro graphics card.
\vspace{-0.3cm}
\section{Results and discussion}
\label{sec:result}
\noindent We conducted our experiments on GPUs via the caffe framework, 
\cite{jia2014caffe}. We present our results in table \ref{tab:res} where 
we averaged the metrics over all left out patients. We also added, for 
illustration purposes \ref{fig:prediction}, image predictions and 
associated 
probability maps to evaluate some differences between the classifiers. 
We notice that the simpler net, PangNet seems to have learnt simple 
rules such as color information. As soon as the nuclei are not
homogeneous and dark inside, the method fails ( "holes" in the
cells). On the 
contrary, deeper nets as FCN and DeconvNet have learnt deeper features 
and seem capable of recognizing whole cells. On these probability maps 
we applied our watershed segmentation and we can picture the results in 
the third 
column. By construction, our post-processing 
scheme is very sensitive to noise as we can see in the PangNet 
prediction, many unrelated minimas will lead to highly partitionned cell. 
This phenomenom also appears with the DeconvNet where undesired 
cuts were performed, like for the cell at the bottom left of the image. 
We picked $\lambda = 7$ which seemed to give a nice partitionning of 
the image, this value can be found by cross validation.
\begin{table}
\begin{tabular}{|c|c|c|c|c|}
\hline
  & PangNet & DeconvNet & FCN & Ensemble\\
 \hline
Accuracy  &  0.924 & 0.954 &0.944 & 0.944  \\
IU   &    0.722 &  0.814 & 0.782 & 0.804 \\
Recall     &  0.655  & 0.773 & 0.752 & 0.900 \\
Precision   &  0.814 & 0.864 & 0.823 & 0.741 \\
F1    &       0.676 &     0.805 & 0.763  & 0.802\\
% TP       &       0.744 &     0.856 & 0.855 & 0.876 \\
% TN      &       0.963 &     0.981 &  0.977 & 0.979\\
Performance    &   0.803   & 0.875 & 0.862 & 0.922 \\
% Pixel error  &  0.063 &     0.032 & 0.042 & 0.032\\
\hline
\end{tabular}
\caption{Results}
\label{tab:res}
\end{table}

>>>>

\noindent Many traditional techniques have been proposed for the segmentation of
nuclei in histopathology images, ranging from simple background
subtraction and color threshold techniques 
to much more
sophisticated approaches, such as marked point
processes. Many of these methods have been recently
reviewed \cite{irshad2014methods}.

@@

%Alternative methods 
%have arisen from the achieving Convolutional neural networks.
%\cite{lecun}, 
%in particular strong incentive rose with the very impressive 
%results that they achieved in the pascal voc challenge with 
%\cite{ImageNet}. 
More recently, advances in deep convolutional neural networks (CNN), and in 
particular in their optimization have made them become the state-of-the-
art model for object recognition. 
Indeed, deep neural networks have been used for the localization and classification of nuclei in histopathology data from breast and colon cancer \cite{xu2016stacked,sirinukunwattana2016locality}, but these previously published approaches do no allow for segmentation.

On the other hand, deep CNN have also been used for semantic segmentation in other fields, where \cite{long2015fcn} use "de-convolution layers" and up-sampling in order to directly output binary images. The rationale of our proposed method was therefore to learn segmentation directly from annotated samples with this or similar architectures.

%identify and precisely 
%locate objects within a picture. Different architectures arising from 
%different intuition are also possible and have been applied in this paper.

Learning based methods rely on annotated data sets. But for the
problem of nuclei segmentation in histopathology images, there are
only few manually segmented images freely available. 
In \cite{Drelie08-298}, a partially annotated data set has been
released. However, this data set does not cover the morphological
variability typically encountered in real histopathology data.

\begin{figure}[htb]
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_1}}
%  \vspace{1.5cm}
%  \centerline{Sample from patient 498959}\medskip
  \centerline{Sample from patient 1}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_1}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
%\begin{minipage}[b]{.48\linewidth}
 % \centering
%  \centerline{\includegraphics[width=4.0cm]{RGB_2}}
%  \vspace{1.5cm}
%  \centerline{Sample from patient 581910}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{0.48\linewidth}
%  \centering
%  \centerline{\includegraphics[width=4.0cm]{GT_2}}
%  \vspace{1.5cm}
%  \centerline{Associated ground truth}\medskip
%\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_3}}
%  \vspace{1.5cm}
 % \centerline{Sample from patient 581910}\medskip
  \centerline{Sample from patient 2}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_3}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
\caption{Random annotated samples from the dataset}
\label{fig:annotation}
%
\end{figure}

@@

%% Validation scheme
We have our training set $(A_l,B_l)_{l \in [1, N]}$ where N is equal to 
33, however each element $A_l$ belongs to a certain patient and 
several elements $A_l$ can belong to the same patient. As we are 
dealing with histopathology images, it is known that samples can widely 
vary from one patient to the other. We thus validate our model by a 
leave-one-patient-out scheme: for a 
given set of hyper parameters, we 
train our model on every patient except one that is used for validation. 
Our final score is averaged over all patients.
Several metrics give a detailed assessment of the model quality: accuracy, intersection over union (IU),
F1 score and a performance score which is the mean between 
true positive rate and true negative rate.

@@

%
\caption{Prediction via different classifiers of a random sample on the (same) left out patient}
\label{fig:prediction}
%
\end{figure}

@@

%\vspace{-.2cm}
\section{Different architectures}
\label{sec:results}
\noindent We experience with 3 known arhitectures in semantic segmentation, named PangNet, Fully Convolutionnal Net (FCN) and DeconvNet. The most basic architecture, PangNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map, for more information please refer to \cite{pang2010cell}. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
architecture has the advantage of re-using a classical deep learning 
architecture with additional upsampling and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture\cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layers as one intends to learn 
the proper upsampling through repeated deconvolution and convolution 
layers. This model can also be fine tuned\cite{noh2015learning}. We also did an
ensemble classifier of these two previous nets as suggested in 
\cite{noh2015learning} but did not include any illustration. Processing 
one image of size $224 \times 224$ pixels took on average 0.91 secondes for 
PangNet, which is the smallest network, and it took on average 3.38 secondes for DeconvNet, the 
largest network. These times do not depend on the number of cells. These computation 
times were estimated on a K620 Quadro graphics card.
\vspace{-0.3cm}
\section{Results and discussion}
\label{sec:result}
\noindent We conducted our experiments on GPUs via the caffe framework, 
\cite{jia2014caffe}. We present our results in table \ref{tab:res} where 
we averaged the metrics over all left out patients. We also added, for 
illustration purposes \ref{fig:prediction}, image predictions and 
associated 
probability maps to evaluate some differences between the classifiers. 
We notice that the simpler net, PangNet seems to have learnt simple 
rules such as color information. As soon as the nuclei are not
homogeneous and dark inside, the method fails ( "holes" in the
cells). On the 
contrary, deeper nets as FCN and DeconvNet have learnt deeper features 
and seem capable of recognizing whole cells. On these probability maps 
we applied our postprocessing scheme and we can picture the results in 
the third 
column. By construction, the postprocessing relies on a smooth posterior probability as we can see in the PangNet 
prediction: many unrelated minimas will lead to oversegmentation. 
This phenomenom also appears with the DeconvNet where undesired 
cuts were performed, like for the cell at the bottom left of the image. 
We picked $\lambda = 7$ which seemed to give a nice partitionning of 
the image, this value can also be found by cross validation.
\begin{table}
\begin{tabular}{|c|c|c|c|c|}
\hline
  & PangNet & DeconvNet & FCN & Ensemble\\
 \hline
Accuracy  &  0.924 & 0.954 &0.944 & 0.944  \\
IU   &    0.722 &  0.814 & 0.782 & 0.804 \\
Recall     &  0.655  & 0.773 & 0.752 & 0.900 \\
Precision   &  0.814 & 0.864 & 0.823 & 0.741 \\
F1    &       0.676 &     0.805 & 0.763  & 0.802\\
% TP       &       0.744 &     0.856 & 0.855 & 0.876 \\
% TN      &       0.963 &     0.981 &  0.977 & 0.979\\
Performance    &   0.803   & 0.875 & 0.862 & 0.922 \\
% Pixel error  &  0.063 &     0.032 & 0.042 & 0.032\\
\hline
\end{tabular}
\caption{Results}
\label{tab:res}
\end{table}
